---
title: "Project 2"
author: "Kolton Wiebusch"
date: "10/14/2020"
output: 
      github_document:
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, cache = TRUE)
```

# Required Packages
```{r packages}
library(tidyverse)
library(caret)
library(knitr)
library(rpart)
```

# Introduction

## Purpose: 
Use bike sharing data to analyze and create models predicting the cnt response variable. 

## Methods:
Load data in and select important variables only. Filter results for each day of the week and create training and test sets from that data. Produce some summary statistics and plots about the training data to explore it. Fit a regression tree based model and a boosted tree model on the data, then apply the fits on the test sets. 

```{r data}
#Read in the data
bike <- read_csv("day.csv")

#Getting an idea for what predictors have the best
summary(lm(cnt ~ . - casual - registered, data = bike))

#Recreating the bike dataset with variables of interest
bike %>% select(season, holiday, weekday, workingday, weathersit, atemp, hum, windspeed, cnt) -> bike
```

## Data: 
This data contains the daily count of rental bikes between 2011 and 2012 in Capital bike share system with the corresponding weather and seasonal information.

## Variables: 
After looking to see what variables had high p-values in a linear model when predicting the response of cnt, as well as removing casual and registered due to the instructions, I narrowed it down to 9. Cnt represents the count of rental bikes daily. Season can be broken down into 1 = winter, 2 = spring, 3 = summer, 4 = fall. Holiday returns 1 if the day is a holiday. Weekday has a numeric value representing each day of the week, from 0 = Sunday to 6 = Saturday. Workingday returns 1 if the day is neither a weekend nor holiday. Weathersit has 4 levels increasing in intensity: 1 - clear or few clouds, 2 - Mist or cloudy, 3 -light snow or light rain with some thunderstorms and scattered clouds, 4 - Heavy Rain/Sleet/Snow/Thunderstorms (More severe). Atemp is a derived value of normalized feeling temperature. Hum is normalized humidity values divided to 100 (max). Windspeed is normalized wind speed values divided to 67 (max).  


# Data Split
```{r dataSplit}
#Filtering the data by weekday
bike %>% filter(weekday == 0) -> newbike

#Splitting the data 70/30
set.seed(23)

train <- sample(1:nrow(newbike), size = nrow(newbike)*0.7)
test <- dplyr::setdiff(1:nrow(newbike), train)

bikeTrain <- newbike[train, ]
bikeTest <- newbike[test, ]
```


# Summarizations

These summary statistics give several looks at the response variable cnt, or the count of rented bikes on this specific weekday for the training data. The first one shows basic summary statistics for the count. The second breaks the average and standard deviation of daily count down by each season (1 = winter, 2 = spring, 3 = summer, 4 = fall). The third breaks the average of daily count down by weather situation on this weekday (1 = clear, 2 = cloudy, 3 = light precipitation, 4 = heavy or severe precipitation).

```{r summarizations}
#Summary stats of the response variable cnt
summary(bikeTrain$cnt)

#Summary stats of mean and standard deviation of count of rented bikes by season on this weekday
bikeTrain %>% group_by(season) %>% summarise(avg = mean(cnt), sd = sd(cnt))

#Summary stats of mean of count of rented bikes by weather situation on this weekday
bikeTrain %>% group_by(weathersit) %>% summarise(avg = mean(cnt))
```

# Plots

These scatter plots give a visual of how the response variable cnt, or count of the rented bikes on this specific weekday for the training data, correlates with several other continuous predictors. Overall, we can see that temperature has the strongest correlation with the count; as temperature increases, the number of bikes rented increases. The correlations with humidity and wind speed are not quite as strong, but it does appear that as wind speed increases, generally, the number of bikes rented decreases somewhat. 

```{r plots}
wind <- ggplot(bikeTrain, aes(x = windspeed, y = cnt))
wind + geom_point() + labs(title = "Count vs Wind Speed", x = "Wind Speed (Percentage of Max of 67 MPH)", y = "Bike Rentals") + geom_text(x = .35, y = 7500, size = 3, label = paste0("Correlation = ", round(cor(bikeTrain$windspeed, bikeTrain$cnt), 2)), col = "red")

temp <- ggplot(bikeTrain, aes(x = atemp, y = cnt))
temp + geom_point() + labs(title = "Count vs Temperature", x = "Normalized Temperature Proportion in Celsius", y = "Bike Rentals") + geom_text(x = .2, y = 7500, size = 3, label = paste0("Correlation = ", round(cor(bikeTrain$atemp, bikeTrain$cnt), 2)), col = "red")

humidity <- ggplot(bikeTrain, aes(x = hum, y = cnt))
humidity + geom_point() + labs(title = "Count vs Humidity", x = "Normalized Humidity Values (Percentage of Max of 100)", y = "Bike Rentals") + geom_text(x = .35, y = 7500, size = 3, label = paste0("Correlation = ", round(cor(bikeTrain$hum, bikeTrain$cnt), 2)), col = "red")
```

# Modeling





